### 2013-05-29 docs/irmwriteup.tex
writeup of plan

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[citestyle=numeric]{biblatex}
\addbibresource{connectomics motif.bib}

\begin{document}

\title{Nonparametric stochastic blockmodels with spatial structure}
\author{Eric Jonas}

\maketitle

\section{overview}

Classic stochastic blockmodels posit the existence of a latent class
which sets the probability of a relation holding between any
collection of entities in a domain. That is, if $e^1_i$ is an entity
in domain $T1$ and in latent class $m$ and $e^2_j$ is an entity in
domain $T2$ in latent class $n$, then 
\begin{equation}
\operatorname{Relation}(e^1_i, e^2_j) \sim F(\cdot | \eta_{mn})
\end{equation}

Kemp \parencite{Kemp2006a} extends stochastic blockmodels to the nonparametric setting.

For our application we're interested in the directed graph of synapse
connectivity between all pairs of cells. Thus our relation is
``synapses on'' and is defined on $T1\times T1$. We want to determine
the underlying latent class of each cell (``cell type''), as well as
the cell-type-specific connectivity. For this sort of binary
(connected-or-not) data, we'd traditionally use a conjugate
beta-Bernoulli likelihood -- $\eta_{mn}$ is the probability of a
connection between cells of class $m$ and $n$.

In neural systems, we know that we have cell-type-specific
connectivity, but we also know that geometry matters tremendously. In
the retina, certain cell-types tessellate the entire retina, and only
synapse with a subset of neighboring cells. This spatially-selective
connectivity runs very counter to the blockmodel assumptions of latent class
being all that matters.

Based on an off-handed comment about a ``discriminative blockmodel''
that Kevin Murphy, we can extend the connectivity with

\begin{equation}
P(e_i, e_j) \sim \operatorname{Bernoulli}(w_{ab}\cdot f(e_i, e_j)
\end{equation}

where the goal is to have $w_{ab}$ be an arbitrary latent-class weight
matrix that is applied to an arbitrary discriminative function between
additional parameters of entities $e_i$ and $e_j$. That is, $f$ can be
defined on additional per-entity attributes, like spatial location.

But of course, we don't have to be so limiting -- we can in fact
learn an arbitrary class-type-specific function $f_{ab}(e_i, e_j)$. Right
now I'm planning Euclidian distance between $e_i$ and $e_j$ as well as learning
the $\lambda_{ab}$. That is, the probability of a connection between
cells $e_i$ and $e_j$ is now

\begin{equation}
p(e_i, e_j) \sim \operatorname{Bernoulli}\Big(p_{ab}\cdot \operatorname{Exp}(d(e_i, e_j) | \lambda_{a, b})\Big)
\end{equation}

\begin{itemize}
\item The choice of an exponential distribution is entirely arbitrary
  -- it restricts us to assuming that cells always prefer closer
  neighbors than far away ones. Thus we'll probbably eventually want
  to use something like a two-parameter gamma to enable ``cell class A
  only synapses on cell class B when it is more than 40 $\mu m$ away
  and less than $200 \mu m$ away.''
\item We could even go crazy and stick a DP in there and learn
  whole-hog the functional form of $f_{ab}$. But that sounds
  like overkill right now. 
\end{itemize}

\section{Inference}

Inference isn't that bad -- this really just looks like a nonconjugate
IRM! With the original beta-Bernoulli IRM, we integrate out the
$\eta_{ab}$, whereas here we are explicitly tracking the per-block
parameters. We need to stick a prior on them, and then do
non-conjugate inference, but thankfully other
people \parencite{Neal2000} have already done a passable job
making MCMC in nonconjugate models a bit easier.

As we get ``real'' large data, we'll want to certainly explore
more efficient inference approaches, but we can fight that fight
when we come to it.

Other things: we're estimating the latent class number, should we be
using a mixture of finite mixtures? The DP is inconsistent
\parencite{Miller2012} but minor tweaks allow us to use nearly-identical
inference and more robustly learn the number of latent categories.

\printbibliography

\end{document}