### 2013-12-11 docs/paper.tex
Latest version of text

%\documentclass{nature}
\documentclass{article}
\usepackage[english,american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}

\renewcommand{\vec}[1]{\mathbf{#1}}

\addbibresource{connectomics motif.bib}
\addbibresource{connectomics motif-relational models.bib}

\DeclareGraphicsRule{.ai}{pdf}{.ai}{}

\title{Discovering cell type motifs in connectomics data}
\author{Eric Jonas, Konrad Kording}

\begin{document}
\maketitle

\todo{Citations}
\todo{Log score plots for real data} 
\todo{Hierarhcical structure?}
% \begin{affiliations}
%   \item Something Something
%   \item Something Something
% \end{affiliations}

\begin{abstract}
  New techniques produce massive data about neural connectivity,
  necessitating new analysis methods to discover underlying
  structure. There is strong evidence for microcircuitry, where the
  probability of connections depends on the types of pre- and
  post-synaptic neurons and their distance. Here we developed a
  nonparametric Bayesian technique that identifies neuron type and circuitry
  patterns using only this connection data. We show that the approach recovers
  known neuron types in the retina, reveals interesting structure in
  the nervous system of c. elegans, and automatically discovers the
  structure of microprocessors. Our approach is a first step towards
  condensing connectomics data into meaningful, human readable
  structure.
\end{abstract}

\section*{Introduction}
Emerging techniques \autocite{Lichtman, Zador, Denk} promise to
quantify the location of each neuron within a volume of interest,
along with its connections to all other neurons. Far exceeding the
capacity of neuroanatomists to trace small circuits, this leads to
enormous datasets that quantify aspects of the nervous system. The impact of
this kind of high-throughput transition has been seen before -- 
rise of sequencing techniques necessitated the
development of novel computational methods to understand genomic
structure, ushering in an era of bioinformatics as an independent
discipline \autocite{}.

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another. In fact, the idea of well defined local connectivity patterns
has been prominent in many areas, ranging from sensory (e.g. retina,
\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal
cord) \autocite{}. Indeed, such repeated patterns also exists in
human-made computing circuits. It remains an important challenge to
develop algorithms to use anatomical data, e.g. connectomics, to back
out underlying connectivity motifs.

The discovery of structure is a crucial aspect of network
science. Early approaches focused on global graph properties, such as
the types of scaling present in the network \autocite
{WattsStrogatz1998} .  While this approach leads to an understanding
of global network, more recent work aims at identifying very small-scale
repeat patterns, or “motifs” in networks\autocite{Milo2002}

The discovery of structure in probabilistic graphs is a crucial aspect
in machine learning. Commonly used algorithms include spectral methods
\autocite{},   community-based-detection   methods  \autocite{},   and
stochastic block  models \autocite{NowickiSnijdders2001}.  While these
approaches powerfully  incorporate the probabilistic nature of neural
connections \autocite{} they do not generally consider the known fact
that  the  probability  of  connections  between  neurons  depends  on
distance. Combining such probabilistic models with a spatial component
to model distance promises to  improve the detection of microcircuitry
from neural data.

When it comes to neuroscience and other computing systems, we expect
patterns of connectivity much more compelx than traditional motifs,
exhibiting a strong spatial dependence arising from the complex
genetic, chemical, and activity-based neural development process.

Here we describe a Bayesian non-parametric model that can discover
circuit structure automatically from connectomics data: both the cell
types and their spatial patterns of interconnection. We apply it to
three computing systems: the mouse retina connectome \autocite{}, the
c. elegans connectome \autocite{}, and a ``connectome'' of a classical
microprocessor. Note we do not explicitly consider cellular
morphology, absolute position (cell layer), or any chemical or genetic
information. Yet in all cases, we discover cell-type to cell-type
connectivity that depends on spatial distance and these connectivity
functions depend on the two involved cell types. Comparing the cell
types discovered by the algorithms with those obtained by anatomists
based on orthogonal information reveals a high degree of agreement. We
present a scalable approach to infer microcircuitry from connectivity
data.

\begin{figure}
  \centering 
    \centerline{\includegraphics[width=183mm]{overview.ai}}
  \caption{An overview of our method. a. Input data is
an adjacency matrix indicating whether cell $i$ (row)
synapses on cell $j$ (column). The location of each cell in space is known. 
b. Our algorithm discovers hidden cell types in this connectivity data
by assuming all cells of a type share a distance-dependent connectivity 
profile, c.), with cells of other types. This results in a clustering 
of the cells by those hidden types. b.) shows the cell adjacency
matrix with cells of the same type grouped together. d.) shows
the learned probability of connection between our different types
at various distances, illustrating short-range and long-range
connectivtiy patterns.}

\end{figure}

\section{Results}
We use a Bayesian structure discovery approach that is based on
assumptions about the way positions and cell types affect the
probability of connections between cells, a so-called generative
model. The probability of connections between pairs of neurons
decreases with distance, but how high the baseline probability of
connection is and how quickly it decays with distance depends on the
two cell-types. We combine annealing with Markov Chain Monte Carlo to
perform simultaneously joint posterior inference over the structure,
parameters, and hyperparameters of the model (see Methods). This
procedure yields a division of all neurons into cell-types along with
the distance and type dependent connectivity probabilities. These
results can then be compared with the results of previous anatomical
studies that have used other aspects of neurons, such as chemical
properties and cell-morphology to divide neurons into types.

Before we can use a probabilistic model to analyze real data we first
need to assess if the model works properly. We thus started out by
simulating data for which we know the correct structure and comparing
the estimated structure based on the algorithm (see methods) with the
one we used for simulation. We find that the model does a good job of
recovering the structure we used to generate the data (Fig
1A). Indeed, it does far better than a simpler and well established
model (stochastic block) that assumes that there is no distance
dependence to the connectivity. The model converges relatively
quickly, which is enabled by xxx cool tricks we used to speed it
up. It appears that the model behaves the way it should and allows
application to large datasets.

\begin{figure}
  \centering 
  \centerline{\includegraphics{synthetic.ai}}
  \caption{Correct recovery of true numbers of hidden classes
in synthetic data. a.) Other models overcluster the number
of classes as they fail to take distance into account, whereas
our model finds close to the true number of classes. b.) ARI, a measure
of ``correct clustering'', for
different true class counts, between our model and a traditional
nonparametric stochastic block model. c.) Multiple Markov chains iterateively
converge towards a highly-probable solution. }

\end{figure}

\subsubsection{Mouse Retina}
The retina of the mouse \autocite{Masland2001} is an example where we
should expect to have connectivity patterns that are well approximated
by our generative model. It is known that there are multiple classes
of cells that can be grouped into ganglion cells that transmit
information to the rest of the brain, bipolar cells that connect
between different cells, and amacrine cells that feed into the
ganglion cells (cite some review). Recent research \autocite{Helmstaedter2013} has
produced a large dataset containing both the types of cells from
orthogonal approaches, and also the connectivity matrix between all
reconstructed cells. We want to know to which level we can classify
neurons into types based on this connectivity only.

The algorithm took 8 hours to converge to a locally optimal solution,
dividing neurons into a set of cell types (Fig 2a). For each pair of
neurons there is a specific distance dependent connection probability
(Fig 2b), which is well approximated by the model fit. Moreover, each
type of cell is rather isotropically distributed across space (Fig 2c)
as should be expected for true cell types. The algorithm is thus able
to extract meaningful structure from the data.

Comparing the results of the algorithm to other information sources
allows evaluating the quality of the type determination. We find that
the types tend to reflect the known laminar distribution in the retina
(Fig 2a, outer ring), which is exciting as the algorithm did not
actually use this type of information. Moreover, we can compare the
results from our algorithm with cell-type determinations by
professional anatomists, who distinguish between 71 types of
neurons. The algorithm yields a separation of neurons into a smaller
number of types that are, however, highly correlated with those of the
anatomists.

Our types closely reflect the (human-determined) segmentation of cells
into retinal ganglion, narrow amacrine, medium/wide amacrine, and
bipolar cells (Figure 2a, outermost ring). [TODO: Do we actually
assign any of the unknown cells correctly? are types 72-79 “differet”
or simply unknown?]

\begin{figure}
  \centering 
  \centerline{\includegraphics{mouseretina.ai}}
  \caption{Discovering cell classes in the mouse retina connectome. 
a.) Input adajacency data for 950 cells for which soma positions were known. b. clsutered adjacency matrix. c.) connectivity diagram showing our clusters, as
well as the cell depth and known cell type of each cell. Our model agrees
with the cell types traditionally identified by anatomists, and identifies 
laminar-specific connectivity patterns. d.) Connectivity between our
clusters as a function of distance -- the cluster consisting primarially of
retinal ganglion cells (lower left-center) exhibits the expected near and
far connectivity.}
\end{figure}

\subsection{C. elegans}

The roundworm Caenorhabditis elegans is a model system in
developmental neuroscience\autocite{White1986}, with the location and
connectivity of each of 302 neurons deterministically known. This
incredible cross-organism consistency led to early mapping of the
c. elegans chemical and electrical (gap junction) connectome. Unlike
the retina, only the motor neurons in c. elegans exhibit regular
distribution in space, being distributed along the body axis. Most
interneurons are concentrated in various ganglia that project
throughout the entire animal, and the sensory neurons are primarily
located in a small number of very anteriior ganglia.

Using both the chemical and electrical connectivity (see methods), we
determined the underlying clusters explained by connectivity and
distance (fig ba). A superficial inspection of the results shows
clustering into groups consisting roughly homogeneously of motor
neurons, sensory neurons, and interneurons. Closer examination reveals
agreement with the classifications originally outlined by Brenner in
1986.

We identify cell types that reflect the known motor/non-motor neuron
classes, even though this system lacks the strong repeat
microcircuitry our model was designed for.  Motorneuron types AS, DA,
and VA, all exclusively postsynaptic, are clustered together, as are
motorneuron types VD and DD. Traditional types VC, DB, and VB also
mostly share a cluster. Various head motor neurons, including types
SMD, RMD, and RIM are clustered together. Interneurons with known
anatomically-distinct connectivity patterns, such as RIA (2 cells) or
AVA (2 cells) are clustered into pure types. The algorithm even
correctly places the single-cell type DVB by itself. Note our
clustering does not perfectly reflect known divisions -- several
combinations of head and sensory neurons are combined, and a
difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
grouped together. Our identified cell types thus reflect a
``coarsening'' of known types, based entirely on connectivity and
distance information, evne when the organism exhibits substantially
less spatial regularity than the retina.

\begin{figure}
  \centering 
  \centerline{\includegraphics{celegans.ai}}
  \caption{Discovering cell classes in the \textit{c. elegans}
    connectome. a.) Initial cell connectivity matrix, red are
    (directional) chemical synapses, blue are electrical
    synapses. Point size is proportaionl to synapse count between
    those cells. b.) Clustered adjaceny matrix showing identified
    clusters. c.) Inter-cluster connectivity, showing the position of
    each cell along the body axis as well as the ``true'' cell type --
    both coarse (motor, sensory, and interneuron) as well as fine
    (small text labels). }
\end{figure}

\subsection{Microprocessor}

To show the applicability of our method to other connectome-style
datasets, we obtained the spatial location and interconnectivity of
the transistors in a classic microprocessor, the MOS Technology 6502
(used in the Apple II) \autocite{datasource}. Computer architects use
common patterns of transistors when designing circuits, with each
transistor having a ``type'' in the circuit. We identified a region of
the processor with complex but known structure containing the primary
8-bit registers X, Y, and S (figure).

Our algorithm identifies areas of spatial homogeneity that mirror the
known structure in the underlying architectural circuit, segementing
transistor types recognizable to computer architects. Using the
original schematics, we see that one identified type contains the
``clocked'' transistors, which retain digital state. Two other types
contain transistors with pins C1 or C2 connected to ground, mostly
serving as inverters.  An additional identified type controls the
behavior of the three registers of interest (X, Y, and S) with respect
to the SB data bus, either allowing them to latch or drive data from
the bus. The repeat patterns of spatial connectivity are visible in
figure~\ref{fig:microprocessor}a, showing the man-made horizontal and
vertical layout of the same types of transistors.

\begin{figure}
  \centering 
  \centerline{\includegraphics{mos6502.ai}}
  \caption{Microprocessor data}
\end{figure}

\section{Discussion}
We have presented a machine learning technique that allows cell types
and microcircuitry to be discovered using only distance and
connectivity data. We have shown its applicability to a newly
published connectomics dataset from the mouse retina, the classical
connectivity of the worm c. elegans, and the connectivity of a
historical man-made microprocessor. We have found that distance and
connectivity alone are sufficient to discover cell types and
microcircuitry that were known to exist in the systems based on decades
of previous research which had utilized extensive additional
information.

For probabilistic models like ours, no known solution exists to
exactly find the most probable parsing of the neurons into cell-types
and connectivity patterns . We employ a collection of Markov-chain
Monte carlo techniques [reference methods] but while different
initializations converge to similar ultimate values, we can never
realistically obtain the global optimum. There are a broad range of
techniques that may offer better approximations to the global optimum
\autocite{Wanga2012,FritzJonathan} and future work could adapt them to
the problem we addressed here.

Our inference becomes slower as the amount of data increases. Our
algorithm required 30 hours for 1000 neurons and could take months for
substantially longer datasets. Scaling this class of probabilistic
model is an active area of research, and recent results in both
variational methods \autocite{Hoffmann2013} and spectral learning
  \autocite{Anandkumar2012} promise substantial speed increases.

  Larger datasets will generally allow algorithms to distinguish more
  distinct types (cite supplemental data) and more data should
  algorithms to get closer to the results from anatomy. Moreover, in
  general, for such problems precision increases with the size of the
  dataset and the cells that we have are not sufficient to
  statistically distinguish all the cell types known in anatomy
  (71xxx). Still, using only connectivity and distance it is possible
  to meaningfully divide neurons into types.

Our small collection of hand-selected distance-dependent likelihood
functions are clearly non-exhaustive, and assume monotonicity
of connectivity probability -- for a given class, closer cells
are never less-likely to connect. This is known to be insufficient
for various neural systems. Future models could incorporate
a wider variety of likelihood functions, or even learn the global
functional form from the data.

There have existed a a range of previous approaches to the discovery
of neural microcircuitry (cites). These generally involve a great deal
of manual labor and ad-hoc determination of what constitutes a “type”
of cell -- to this day there are disagreements in the literature as to
the “true” types in the mammalian retina. Much as phylogenomics has
changed our understanding of animal ontologies, modern large scale
data will allow the efficient unbiased discovery of cell types and
circuits. The sheer amount of available data demands the introduction
of algorithmic approaches.

The development of automatic identification and quantification of cell
type may also provide new phenotypic tools for quantifying the effect
of disease, genetic abnormalities, and genetic
interventions. Particular pathologies and genetic manipulations may
subtly change the number and connectivity of cell types in ways
only visible with algorithmic analysis.

Ultimately, we will need to combine information across different
sources into a joint understanding of neural function. Distinct cell
types differ in morphology, connectivity, transcriptomics, relation to
behavior or stimuli and many other ways. Algorithms based on each type
information can then be combined to synthesize all the available
information from one experiment or even across experiments into a
joint model of brain function.

\section{Methods}
\section{Methods Summary}

% this should clearly describe the inputs

% the outputs

% the model

We take as input a connectivity matrix $R$ defining the connections
between cell $e_i$ and $e_j$, as well as a distance function $d(e_i, e_j)$
representing a (physical) distance between adjacent cells. See the
supplemental material for extension to multiple connectivity
matrices. We assume there exist an unknown number $k$ of latent
(unobserved) cell types, $k \in \{1, 2, 3, \dots\}$, and that
each cell $e_i$ belongs to a single cell type, $c_i = k$. The observed
connectivity between two cells $R(e_i, e_j)$ then depends only on
their latent type and their distance through a link function $f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the latent class, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well as a set of global hyper parameters $\theta$, such that the link function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.

We then jointly infer the maximum a posteriroi (MAP) estimate of the
class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
the global model hyperparameters $\theta$ :

\begin{equation}
  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
\end{equation}

% Methods section 3000 words
% methods summary does not appear in the online text
% methods section can have no figures or tables [grrr]
% todo describe multiple independent relations?

\subsection{Probabilistic Model}

Our model is a extension of infinite stochastic block models
\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints.

\todo{infinite subset relaitonal model has good summaries}

We describe three link functions, ``Logistic-distance Bernoulli'',
``logistic-distance fixed-lambda Bernoulli'', and ``logistic-Distance
Poisson''.

% fixme class component, NOT class
The ``logistic-distance Bernoulli'' link model assumes that, if cell
$e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
and $e_j$ are connected is given by
\begin{eqnarray}
p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
\end{eqnarray}
where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$.

We use a Dirichlet-process prior on cluster assignments, which allows
the number of clusters to be determined automatically. In brief, the
probability of a cell belonging to a cluster is proportional to the
number of datapoints already in that cluster, $m_k$, such that $p(c_i
= k) \propto \frac{m_k}{N + \alpha}$ and the probability of the cell
belonging to a new cluster $k'$ is $p(c_i = k') \propto
\frac{\alpha}{N + \alpha}$.

\subsection{Inference} 
We perform posterior inference via Markov-chain Monte Carlo (MCMC),
annealing on the global likelihood during the traditional burn-in
phase. MCMC transition kernels for different parts of the state space
can be chained together to construct a kernel whose ergodic
distribution is the target ergodic distribution over the entire state space.

Our first transition kernel (``structural'') performs gibbs sampling 
of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
The lack of conjugacy in our likelihood model makes an explicit 
evaluation of the conditional assignment probabilities impossible, 
motivating us to use an auxiliary variable method \autocite{Neal2000}
in which a collection of ephemeral clusters are explicitly represented
for the duration of the Gibbs scan.

We then employ a transition kernel to update the per-component
parameter values $\eta_{mn}$. Conditioned on the assignment vector
$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
individual $\eta_{mn}$ are indepednent. We slice sample \autocite{Neal2003}
each component's parameters, choosing the slice width as a function
of the global hyperparameter range.

The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
to take on a discrete set of possible values. As $\theta$ is often a
tuple of possible values, we explore the cartesian product of all
possible values. We then Gibbs sample, which is always possible in a 
small, finite discrete state space.

We chain these three kernels together, and then globally
anneal on the likelihood from a temperature of $T=64$ down to 
$T=1$ over 300 iterations unless otherwise indicated, and
then run the chain for another $100$ iterations.

To pick the MAP, we run up to $100$ simultaneous chains in parallel,
iniitalized from different random initial points in the statespace,
and then pick the chain with the highest log likelihood. Early
experiments attempted to use the full ensemble of chains as an
estimate of the posterior but were abandoned due to the difficulty of
summarizing the posterior over such a complex state space, whcih is an
ongoing area of research in the probabilistic modeling community
\autocite{}.

\subsection{Validation}

\subsection {Mouse Retina}
Dense serial electron microscopy of a $VOLUMNE$ in the mouse retina by
\autocite{Helmstaedter2013} yielded a listing of places where neurons
come into contact. There were over 1000 cells originally, and selected
the $950$ for which the location of the soma could be reconstruted
from the provided cell plots (soma locations were not provided by the
study's authors in machine-readble form). Ultimately
this left a matrix between the total synapse-like contact area between
all pairs of 950 cells. Area was thresholded at $0.5\mu m$, determined
by hand, to yield a 950 $\times$ 950 entry matrix that served as input
to our algorithm. We measured the distance between cells using
the reconstructed soma centers, and used the Logistic-Distance
spatial relation. Hyperprior griddings are shown in \ref{supplemental}.

\subsection{C. elegans}

We obtained the connectome of c. elegans from \autocite, and isolated
the 279 nonpharyngeal neurons, with a total of 6393 chemical synapses
and 890 gap junctions originally cleaned up in \autocite{Chen2006}. A
cell's position was its distance along the anterior-posterior axis
normalzed between zero and one. We used both networks, the chemical
network as a directed graph and the electrical network as undirected
graph. We use the synapse counts with the logistic-distance poisson
likelihood, scaling the counts by 4.0 to compensate for 
the Poisson's overdispersion.

\subsection{Microprocessor}
We extracted the connection graph for the transistors in the MOS6502
\autocite{visual6502}. Each transistor has three terminals (gate,
source, drain), but the methods of the original dataset were unable to
consistently resolve which of the C1 and C2 terminals were source and
drain, leading to ambiguity in our encoding. We identified a region
consisting of consisting of three registers X, Y, and S via visual
inspection and focused our efforts there. We created a total of six
connectivity matrices by examining possible terminal parings. One
graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
$e_j$ are connected via pins $g$ and $c_1$.

% \begin{addendum}
%  \item Put acknowledgements here.
%  \item[Competing Interests] The authors declare that they have no
% competing financial interests.
%  \item[Correspondence] Correspondence and requests for materials
% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
% \end{addendum}
\section{Supplemental}
\subsection{Source code}
\todo{Put up website}

All source code and materials for running experiments can be
obtained from the project website, at http://spatialrelation.github.io.

\subsection{Extension to multiple graphs}
The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by simply considering the product of 
\begin{equation}
  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha) 
\end{equation}

\subsection{Microprocessor Netlist annotation}
This is a matrix consisting of the different clusters of nets for 
the registers.

\subsection{Hyperprior grids}

For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80. $p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.

For the c. elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0. We globally set $rate_{min}=0.01$.

For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.

\todo{Explain Adjusted Rand Index}

\printbibliography

\end{document}